On the specific recursive structure of iterated parametricity

### 2025-10-19 paper/paper.tex
paper: progress

@@

\newcommandx{\frametype}[2][1,2]{\prim{frametype}[][#1][#2][][]}
\newcommandx{\layertype}[2][1,2]{\prim{layertype}[][#1][#2][][]}
\newcommandx{\paintingtype}[2][1,2]{\prim{paintingtype}[][#1][#2][][]}
\newcommandx{\restrftype}[2][1,2]{\prim{restrftype}[][#1][#2][][]}
\newcommandx{\restrltype}[2][1,2]{\prim{restrltype}[][#1][#2][][]}
\newcommandx{\restrptype}[2][1,2]{\prim{restrptype}[][#1][#2][][]}
\newcommandx{\cohftype}[2][1,2]{\prim{cohftype}[][#1][#2][][]}
\newcommandx{\cohltype}[2][1,2]{\prim{cohltype}[][#1][#2][][]}
\newcommandx{\cohptype}[2][1,2]{\prim{cohptype}[][#1][#2][][]}

\newcommandx{\deps}[2][1,2]{\prim{deps}[][#1][#2][][]}
\newcommandx{\extradeps}[2][1,2]{\prim{deps}[][#1][#2][][]}
\newcommandx{\depstype}[2][1,2]{\prim{depstype}[][#1][#2][][]}
\newcommandx{\extradepstype}[2][1,2]{\prim{depstype}[][#1][#2][][]}

@@

At this stage, it is convenient to use an abbreviation to combine a $\framep$, a $\painting$ and a $\restrf$ in one block, which we call deps:

\begin{align*}
  \deps[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \framep[n][0,...,p-1] :\frametype[n][0,...,p-1]      \\
            \painting[n][0,...,p-1] :\paintingtype[n][0,...,p-1] \\
            \restrf[n][0,...,p-1]  :\restrftype[n][0,...,p-1]    \\
          \end{array}\right)
\end{align*}
so that $\framep[n][0,...,p]$ has type $\deps[n][0,..,p-1] \rightarrow \frametype[n][0,...,p]$.

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \frametype[n][p,...,n]         &                                              & :         & Type_{l+1}                                                                                   \\
    \frametype[n][]                &                                              & \defeq    & \unittype                                                                                    \\
    \frametype[n][p,...,n]         &                                              & \defeq    & Type_l \times \frametype[n][p+1,...,n-1]                                                     \\
    \\
    \paintingtype[n][p,...,n]      & (\framep[n][p,...,n]:\frametype[n][p,...,n]) & :         & Type_{l+1}                                                                                   \\
    \paintingtype[n][]             & \unitpoint                                   & \defeq    & \unittype                                                                                    \\
    \paintingtype[n][p,...,n](p<n) & (\framep[n][p,...,n])                        & \defeq    & (\framep[n][p] \rightarrow Type_l) \times \paintingtype[n][p+1,...,n](\framep[n][p+1,...,n]) \\
    \\
    \restrftype[n][p,...,n]        &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \framep[n-1][0,...,p-1]:\frametype[n-1][0,...,p-1]                               \\
                \painting[n-1][0,...,p-1]:\paintingtype[n-1][0,...,p-1](\framep[n-1][0,...,p-1]) \\
                \framep[n-1][p,...,n-1]:\frametype[n-1][p,...,n-1]                               \\
                \painting[n-1][p,...,n-1]:\paintingtype[n-1][p,...,n-1](\framep[n-1][p,...,n-1]) \\
                \restrf[n][0,...,p-1]:\restrftype[n][0,...,p-1]
                \left(\begin{array}{l}
                    \framep[n-1][0,...,p-1]   \\
                    \painting[n-1][0,...,p-1] \\
                  \end{array}\right)                                                  \\
              \end{array}\right) : Type_{l}
    }                                                                                                                                                                                        \\
    \restrftype[n][]               &
    \left(\begin{array}{l}
              \framep[n-1][0,...,n-1]   \\
              \painting[n-1][0,...,n-1] \\
              \unitpoint                \\
              \unitpoint                \\
              \restrp[n][0,...,n-1]     \\
            \end{array}\right)   & \defeq                                       & \unittype                                                                                                    \\
    \restrftype[n][p,...,n-1]      &
    \left(\begin{array}{l}
              \framep[n-1][0,...,p-1]   \\
              \painting[n-1][0,...,p-1] \\
              \framep[n-1][p,...,n-1]   \\
              \painting[n-1][p,...,n-1] \\
              \restrf[n][0,...,p-1]     \\
            \end{array}\right)
                                   & \defeq                                       &
    \begin{array}{c}
      \framep[n][p]
      \left(\begin{array}{l}
                \framep[n-1][0,...,p-1]   \\
                \painting[n-1][0,...,p-1] \\
                \restrf[n][0,...,p-1]     \\
              \end{array}\right)
      \rightarrow \framep[n-1][p] \\
      \times                      \\
      \restrftype[n][p+1,...,n]
      \left(\begin{array}{l}
                \framep[n-1][0,...,p]       \\
                \painting[n-1][0,...,p]     \\
                \framep[n-1][p+1,...,n-1]   \\
                \painting[n-1][p+1,...,n-1] \\
                \restrf[n][0,...,p]         \\
              \end{array}\right)
    \end{array}
  \end{array}
\end{equation*}

@@

\begin{align*}
  \extradeps[n][p,..,n-1] (\deps[n][0,...,p-1]) \defeq
  \left(\begin{array}{l}
            \framep[n][p,...,n-1] :\frametype[n][p,...,n-1]      \\
            \painting[n][p,...,n-1] :\paintingtype[n][p,...,n-1] \\
            \restrf[n][p,...,n-1]  :\restrftype[n][p,...,n-1]    \\
          \end{array}\right)
\end{align*}

This allows to define $\painting[n][p]$ formally:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][p]              &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \deps[n-1][0,...,p-1]:\depstype[n-1][0,...,p-1]                                \\
                \framep[n-1][p,...,n-1]:\frametype[n-1][p,...,n-1]                             \\
                \painting[n-1][p,...,n-1]:\paintingtype[n-1][p,...,n-1](\deps[n-1][p,...,n-1]) \\
                \left(\begin{array}{l}
                    \framep[n-1][0,...,p-1]   \\
                    \painting[n-1][0,...,p-1] \\
                  \end{array}\right)                                                \\
                \restrf[n][p,...,n-1]:\restrftype[n][p,...,n-1]
                \left(\begin{array}{l}
                    \deps[n-1][0,...,p-1]     \\
                    \framep[n-1][p,...,n-2]   \\
                    \painting[n-1][p,...,n-2] \\
                  \end{array}\right)                                                \\
                E:\framep[n-1][n-1]\rightarrow Type_l                                          \\
              \end{array}\right) : \framep[n-1][n-1]\rightarrow Type_l
    }                                                            \\
    \painting[n][n]              &
    \left(\begin{array}{l}
              \framep[n-1][0,...,n-1]   \\
              \painting[n-1][0,...,n-1] \\
              \unitpoint                \\
              \unitpoint                \\
              \restrp[n][0,...,n-1]     \\
              \unitpoint                \\
              E                         \\
            \end{array}\right) & \defeq & E                        \\
    \painting[n][p](p<n)         &
    \left(\begin{array}{l}
              \framep[n-1][0,...,p-1]   \\
              \painting[n-1][0,...,p-1] \\
              \framep[n-1][p,...,n-1]   \\
              \painting[n-1][p,...,n-1] \\
              \restrf[n][0,...,p-1]     \\
              \restrf[n][p,...,n-1]     \\
              E                         \\
            \end{array}\right) & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:(\Pi q\leq n.\,\Pi \omega.\
              \painting[n-1][0,...,p](\restrf[n][p](d))). \\
              \painting[n][p+1,...,n-1]
              \left(\begin{array}{l}
                  \framep[n-1][0,...,p]       \\
                  \painting[n-1][0,...,p]     \\
                  \framep[n-1][p+1,...,n-1]   \\
                  \painting[n-1][p+1,...,n-1] \\
                  \restrf[n][0,...,p]         \\
                  \restrf[n][p+1,...,n-1]     \\
                  E                           \\
                \end{array}\right)(d,l)           \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

@@

We can now define the types of sequences of $\framep$ and $\painting$ as follows:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \frametype[n][0,...,p]    &                                              & :      & Type_{l+1}                                                                                       \\
    \frametype[n][]           &                                              & \defeq & \unittype                                                                                        \\
    \frametype[n][0,...,p]    &                                              & \defeq & \frametype[n+1][0,...,p-1] \times Type_l                                                         \\
    \\
    \paintingtype[n][0,...,p] & (\framep[n][0,...,p]:\frametype[n][0,...,p]) & :      & Type_{l+1}                                                                                       \\
    \paintingtype[n][]        & \unitpoint                                   & \defeq & \unittype                                                                                        \\
    \paintingtype[n][0,...,p] & (\framep[n][0,...,p])                        & \defeq & \paintingtype[n+1][0,...,p-1](\framep[n+1][0,...,p-1]) \times (\framep[n][p] \rightarrow Type_l) \\
  \end{array}
\end{equation*}

@@

                \extradeps[n][p-1,...,n]:\extradepstype[n][0,...,p-2](\deps[n][0,...,p-1]) \\

@@

                \end{array}\right), \restrl[n][p](...)(l)\right)                    \\

@@

@@

>>>>

\newcommandx{\cohtwo}[9][2,3,4,5,6,7,8,9]{
  \ensuremath{\mathsf{\color{chestnut}{coh2}}_{
    \joinargs[\mathsf{\color{indian-yellow}{#1}}][#2][#3][#4][#5][#6]}^{\joinargs[#7][#8]}}
  \setkeys*[bonak]{coh#1}{#9}
}

@@

\newcommandx{\frametype}[2][1,2]{\prim{frametype}[][#1][#2][][]}
\newcommandx{\layertype}[2][1,2]{\prim{layertype}[][#1][#2][][]}
\newcommandx{\paintingtype}[2][1,2]{\prim{paintingtype}[][#1][#2][][]}
\newcommandx{\restrftype}[2][1,2]{\prim{restrftype}[][#1][#2][][]}
\newcommandx{\restrltype}[2][1,2]{\prim{restrltype}[][#1][#2][][]}
\newcommandx{\restrptype}[2][1,2]{\prim{restrptype}[][#1][#2][][]}
\newcommandx{\cohftype}[2][1,2]{\prim{cohftype}[][#1][#2][][]}
\newcommandx{\cohltype}[2][1,2]{\prim{cohltype}[][#1][#2][][]}
\newcommandx{\cohptype}[2][1,2]{\prim{cohptype}[][#1][#2][][]}
\newcommandx{\cohttype}[2][1,2]{\prim{cohttype}[][#1][#2][][]}

\newcommandx{\deps}[2][1,2]{\prim{deps}[][#1][#2][][]}
\newcommandx{\extradeps}[2][1,2]{\prim{deps}[][#1][#2][][]}
\newcommandx{\depscohs}[2][1,2]{\prim{depscohs}[][#1][#2][][]}
\newcommandx{\extradepscohs}[2][1,2]{\prim{depscohs}[][#1][#2][][]}
\newcommandx{\depstype}[2][1,2]{\prim{depstype}[][#1][#2][][]}
\newcommandx{\extradepstype}[2][1,2]{\prim{depstype}[][#1][#2][][]}
\newcommandx{\depscohstype}[2][1,2]{\prim{depscohstype}[][#1][#2][][]}
\newcommandx{\extradepscohstype}[2][1,2]{\prim{depscohstype}[][#1][#2][][]}

\newcommandx{\coht}[6][1,2,3,4,5,6]{\cohtwo{frame}[][#3][#4][#5][#6][#1,#2][][]}

@@

At this stage, it is convenient to use an abbreviation to combine sequences $\framep[n][0,..,p-1]$, $\painting[n][0,...,p-1]$ and $\restrf[n][0,...,p-1]$ in one block, which we call $\deps[n][0,...p-1]$:

\begin{align*}
  \deps[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \framep[n][0,...,p-1] :\frametype[n][0,...,p-1] \\
            \painting[n][0,...,p-1] :\paintingtype[n][0,...,p-1]
            (\framep[n][0,...,p-1])                         \\
            \restrf[n][0,...,p-1]  :\restrftype[n][0,...,p-1]
            \left(\begin{array}{l}
                \framep[n][0,...,p-1]   \\
                \painting[n][0,...,p-1] \\
              \end{array}\right)                   \\
          \end{array}\right)
\end{align*}
so that $\framep[n][0,...,p]$ has type $\deps[n][0,..,p-1] \rightarrow \frametype[n][0,...,p]$. Later, we shall write $\deps.\framep$, $\deps.\painting$, $\deps.\restrf$ for the corresponding projections of a $\deps$.

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \frametype[n][p,...,n]         &                                              & :         & Type_{l+1}                                                                                   \\
    \frametype[n][]                &                                              & \defeq    & \unittype                                                                                    \\
    \frametype[n][p,...,n]         &                                              & \defeq    & Type_l \times \frametype[n][p+1,...,n-1]                                                     \\
    \\
    \paintingtype[n][p,...,n]      & (\framep[n][p,...,n]:\frametype[n][p,...,n]) & :         & Type_{l+1}                                                                                   \\
    \paintingtype[n][]             & \unitpoint                                   & \defeq    & \unittype                                                                                    \\
    \paintingtype[n][p,...,n](p<n) & (\framep[n][p,...,n])                        & \defeq    & (\framep[n][p] \rightarrow Type_l) \times \paintingtype[n][p+1,...,n](\framep[n][p+1,...,n]) \\
    \\
    \restrftype[n][p,...,n]        &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \deps[n-1][0,...,p-1]:\depstype[n-1][0,...,p-1]                                  \\
                \framep[n-1][p,...,n-1]:\frametype[n-1][p,...,n-1]                               \\
                \painting[n-1][p,...,n-1]:\paintingtype[n-1][p,...,n-1](\framep[n-1][p,...,n-1]) \\
              \end{array}\right) : Type_{l}
    }                                                                                                                                                                                        \\
    \restrftype[n][]               &
    \left(\begin{array}{l}
              \deps[n-1][0,...,n-1] \\
              \unitpoint            \\
              \unitpoint            \\
            \end{array}\right)       & \defeq                                       & \unittype                                                                                                \\
    \restrftype[n][p,...,n-1]      &
    \left(\begin{array}{l}
              \deps[n-1][0,...,p-1]     \\
              \framep[n-1][p,...,n-1]   \\
              \painting[n-1][p,...,n-1] \\
            \end{array}\right)
                                   & \defeq                                       &
    \begin{array}{c}
      \framep[n][p]
      \left(\begin{array}{l}
                \deps[n-1][0,...,p-1] \\
              \end{array}\right)
      \rightarrow \framep[n-1][p] \\
      \times                      \\
      \restrftype[n][p+1,...,n]
      \left(\begin{array}{l}
                \deps[n-1][0,...,p]         \\
                \framep[n-1][p+1,...,n-1]   \\
                \painting[n-1][p+1,...,n-1] \\
              \end{array}\right)
    \end{array}
  \end{array}
\end{equation*}

@@

\begin{align*}
  \extradeps[n][p,..,n-1] (\deps[n][0,...,p-1]) \defeq
  \left(\begin{array}{l}
            \framep[n][p,...,n-1] :\frametype[n][p,...,n-1] \\
            \painting[n][p,...,n-1] :\paintingtype[n][p,...,n-1]
            (\framep[n][p,...,n-1])                         \\
            \restrf[n][p,...,n-1]  :\restrftype[n][p,...,n-1]
            \left(\begin{array}{l}
                \deps[n-1][0,...,p]         \\
                \framep[n-1][p+1,...,n-1]   \\
                \painting[n-1][p+1,...,n-1] \\
              \end{array}\right)               \\
          \end{array}\right)
\end{align*}

This allows to define $\painting[n][p]$ formally:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][p]               &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \deps[n-1][0,...,p-1]:\depstype[n-1][0,...,p-1]           \\
                \extradeps[n-1][p,...,n-1]:\extradepstype[n-1][p,...,n-1] \\
                E:\framep[n-1][n-1]\rightarrow Type_l                     \\
              \end{array}\right) : \framep[n-1][n-1]\rightarrow Type_l
    }                                                                            \\
    \painting[n][n]               &
    \left(\begin{array}{l}
              \deps[n-1][0,...,n-1] \\
              \unitpoint            \\
              E                     \\
            \end{array}\right)      & \defeq & E                                   \\
    \painting[n][p](p<n)          &
    \left(\begin{array}{l}
              \deps[n-1][0,...,p-1]      \\
              \extradeps[n-1][p,...,n-1] \\
              E                          \\
            \end{array}\right) & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:(\Pi q\leq n.\,\Pi \omega.\
              \deps.\painting[n-1][0,...,p](\extradeps.\restrf[n][p](d))). \\
              \painting[n][p+1,...,n-1]
              \left(\begin{array}{l}
                  \deps[n-1][0,...,p]          \\
                  \extradeps[n-1][p+1,...,n-1] \\
                  E                            \\
                \end{array}\right)(d,l)                           \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

@@

@@

                \extradeps[n][p-1,...,n]:\extradepstype[n][p-1,...,n](\deps[n][0,...,p-1]) \\

@@

                \end{array}\right), \restrl[n][p]
                  \layer[n+1][p]           \\
                \end{array}\right) (l)\right)                                       \\

@@

    \restrltype[n][p]                                                                                                                                         \\

@@

At this new stage, it is convenient to use an abbreviation to combine sequences $\deps[n][0,..,p-1]$, $\extradeps[n][p,...,n]$, $\restrp[n][0,...,p-1]$ and $\cohf[n][0,...,p-1]$ in one block, which we call $\depscohs[n][0,...p-1]$:

\begin{align*}
  \depscohs[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \deps[n][0,...,p-1] :\depstype[n][0,...,p-1]    \\
            \extradeps[n][0,...,p-1] :\extradepstype[n][p-1,...,n]
            (\deps[n][0,...,p-1])                           \\
            \restrp[n][0,...p-1]: \restrptype[n][0,...,p-1] \\
            \cohf[n][0,...,p-1]: \cohftype[n][0,...,p-1]    \\
          \end{array}\right)
\end{align*}
so that $\restrf[n][0,...,p]$ has type $\depscohs[n][0,..,p-1] \rightarrow \restrftype[n][0,...,p]$. Later, we shall write $\depscohs.\deps$, $\depscohs.\extradeps$, $\depscohs.\restrp$, and $\depscohs.\cohf$ for the corresponding projections of a $\depscohs$.

to do : def of cohptype

Definition of \cohf, using $\UIP$ to solve the dependency in $\coht$.
    \cohf[n][0,...,p]                                                                         &
                \depscohs[n][0,...,p-2]:\depstype[n][0,...,p-2]                                \\
                \extradepscohs[n][p-1,...,n]:\extradepstype[n][p-1,...,n](\deps[n][0,...,p-1]) \\
                \cohp[n][0,...,p-2]:\restrptype[n][0,...,p-1]
                    \depscohs[n][0,...,p-1]      \\
                    \extradepscohs[n][p-1,...,n] \\
                \coht[n][0,...,p]:\cohttype[n][0,...,p]
              \end{array}\right) : \cohftype[n][0,...,p]
    \cohf[n][0]                                                                               &
            \end{array}\right)                                                                    & \defeq & (\unitpoint,\lambda q.\lambda r. \lambda\epsilon. \lambda\omega.\lambda \_.\refl) \\
    \cohf[n][0,...,p+1]                                                                       &
              \depscohs[n][0,...,p]        \\
              \extradepscohs[n][p-1,...,n] \\
              \cohp[n][0,...,p]            \\
              \coht[n][0,...,p]            \\
      \cohf[n+1][0,...,p]
                \depscohs[n+1][0,...,p-1]    \\
                \extradepscohs[n][p-1,...,n] \\
                \cohp[n+1][0,...,p-1]        \\
                \coht[n+1][0,...,p-1]        \\
      (\cohf[n+1][0,...,p], \lambda q.\lambda r.\lambda \epsilon.\lambda \omega.\lambda (\cohf[n][p],\cohl[n][p])). \\
      \left(\cohf[n][0,...,p]
                  \depscohs[n+1][0,...,p-1]    \\
                  \extradepscohs[n][p-1,...,n] \\
                  \cohp[n+1][0,...,p-1]        \\
                  \coht[n+1][0,...,p-1]        \\
                \end{array}\right), \restrl[n][p]
                  \depscohs[n][0,...,p]        \\
                  \extradepscohs[n][p-1,...,n] \\
                  \layer[n+1][p]               \\
                \end{array}\right) (l)\right)                                                                           \\
    \cohttype[n][0,...,p-1]                                                                   &
                \depscohs[n][0,...,p-1]:\depscohstype[n][0,...,p-1]                                    \\
                \extradepscohs[n][p-1,...,n]:\extradepscohstype[n][p-1,...,n](\depscohs[n][0,...,p-1]) \\
                \restrp[n][0,...,p-1]:\restrptype[n][0,...,p-1](\depscohs[n][0,...,p-1])               \\
    \cohttype[n][0,...,0-1]                                                                   &
    \cohttype[n][0,...,(p+1)-1]                                                               &
              \depscohs[n][0,...,p]        \\
              \extradepscohs[n][p-1,...,n] \\
      \Sigma \coht[n+1][0,...,p-1]
      :\cohttype[n+1][0,...,p-1]
                \depscohs[n+1][0,...,p-1]    \\
                \extradepscohs[n][p-1,...,n] \\
      \cohf[n][p][q][\epsilon](\cohf[n+1][p][r][\omega]
                \depscohs[n+1][0,...,p-1]    \\
                \extradepscohs[n][p-1,...,n] \\
      \cohf[n][p][r][\omega](\cohf[n+1][p][q+1][\epsilon]
                \depscohs[n+1][0,...,p-1]    \\
                \extradepscohs[n][p-1,...,n] \\
    \cohl[n][p][q][\epsilon]                                                                  &
              \depscohs[n][0,...,p]:\depscohstype[n][0,...,p]                                        \\
              \extradepscohs[n][p-1,...,n]:\extradepscohstype[n][p-1,...,n](\depscohs[n][0,...,p-1]) \\
              \cohp[n+1][p]:\cohptype[n+1][p]                                                        \\
              \coht[n][0,...,p]:\cohttype[n][0,...,p]                                                \\
      \Pi d.\Pi l. \overrightarrow{\cohf[n][p][r+1][q+1][\epsilon][\omega]}{\restrl[n][p][q][\epsilon](\restrl[n][p][r][\omega](l))} = \\
      \restrl[n][p][r][\omega](\restrl[n][p][q+1][\epsilon](l))
    \cohl[n][p][q][\epsilon][r][\omega]                                                       &
              \depscohs[n][0,...,p]        \\
              \extradepscohs[n][p-1,...,n] \\
              \cohp[n][p]                  \\
    \lambda d. \lambda l. \lambda \theta. (\cohp[n][p][q][\epsilon][r][\omega](l))                                                                                                           \\
